Principles for Open Data Curation: A Case Study with the New York City 311 Service Request Data

### 2024-06-20 manuscript/manuscript.tex
went over abstract and intro

\usepackage[margin = 1in]{geometry}
\usepackage{amsmath}
\allowdisplaybreaks
\usepackage{authblk}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{hyperref}\hypersetup{colorlinks=true, citecolor=blue}
\usepackage{parskip}
\usepackage{textgreek}
\usepackage{graphicx}
\usepackage{booktabs} % For better looking tables
\usepackage{siunitx} % For alignment of numbers
\usepackage{placeins}
\usepackage{caption}  % For custom captions
\usepackage{float}   % Required for the H specifier
\usepackage{adjustbox}
\usepackage[]{lineno}
\linenumbers*[1]

@@

\newcommand{\jy}[1]{\textcolor{orange}{JY: (#1)}}
\newcommand{\yh}[1]{\textcolor{purple}{YH: (#1)}}
\newcommand{\yx}[1]{\textcolor{cyan}{YX: (#1)}}
\let\proglang=\textsf
%% \newcommand{\pkg}[1]{{\fontseries{m}\selectfont #1}}
%% \newcommand\code[2][black]{\textcolor{#1}{\texttt{#2}}}

@@

\begin{abstract}
In the early 21st century, the open data movement began to transform 
societies and governments through principles of transparency, 
innovation, and public engagement. New York City (NYC) has emerged 
as a leader in this movement with the enactment of the Open 
Data Law in 2012, leading to the creation of the NYC Open Data 
portal, which now hosts 2700 datasets from 80 city agencies. This 
resource has proven invaluable for research across various domains, 
including health, urban development, and transportation. The 
success of these initiatives underscores the importance of data 
curation, ensuring the utility and reliability of datasets.

This paper examines the data curation challenges using the 
NYC 311 Service Request (SR) Data as a case study, addressing issues 
of data validity, consistency, and curation efficiency. Based on 
insights from this case study, we propose a set of data curation 
principles tailored for government-released open data. These principles 
aim to enhance data management practices and ensure 
the ongoing utility of open data. The paper concludes with 
actionable suggestions for improving data curation and offers 
general principles for the release of open data.

@@

\noindent
\textit{Keywords:}
	311 Service Request Data;
	Data cleansing;
	Data Consistency;
	Data Curation;
	Data Democratization;
	Data Efficiency;
	Data science;
	Data Validity;
	Government Data;
	NYC Open Data;
	Open Data Movement;
	Open data;
	Public Engagement;
	Quality control;
	Research Data Management;
	Smart Cities;
	Transparency;
\end{abstract}

@@

New York City (NYC) has emerged as a forerunner in the open data
movement, marked by the enactment of the Open Data Law in 2012
\citep{zuiderwijk2014open}. This landmark legislation led to the
creation of the \href{https://opendata.cityofnewyork.us}{NYC Open Data
  portal}, which today hosts an impressive array of 2700 datasets
from 80 different city agencies. This resource has become invaluable
for researchers in various fields as well an enabling local government transparency

In health, datasets have enabled
significant studies on facets of health and healthcare delivery
\citep{cantor2018facets, shankar2021data}. In the realm of urban
development, data has been instrumental in advancing smart city
initiatives \citep{neves2020impacts}. Additionally, transportation
research has benefited greatly from this wealth of data, aiding in the
understanding of urban mobility and infrastructure
\citep{gerte2019understanding}. NYC's Open Data initiative not only
exemplifies commitment to transparency and public engagement but also
illustrates how open data can be a powerful tool in addressing complex
urban challenges. Examples of some of the more popular NYC datasets include:

\begin{itemize}
	\item NYC restaurant health inspection violations
	\item Mapping of car crashes involving pedestrians
	\item Visualization of NYC High School and College enrollment
	\item An interactive dashboard to filter and view charges for jail inmates
	\item Location of City-wide free Internet access points
	\item ...and 220 other complaint types. 
\end{itemize}

Data curation is fundamental in the open data ecosystem, ensuring the
utility and reliability of datasets for diverse applications. Among
the earliest discussions, \citet{witt2009constructing} focus on the
development of data curation profiles tailored to specific contexts,
setting a precedent for targeted data management
strategies. Addressing broader challenges in data sharing and
management, \citet{borgman2012conundrum} highlights the complexities
of research data distribution, emphasizing the need for robust
strategies. This is complemented by the work of \citet{hart2016ten},
who outline essential principles for effective data management,
particularly emphasizing the importance of meticulous curation
practices.

In the realm of collaborative data management,
\citet{beheshti2019datasynapse} underscore the significance of
cooperative environments for managing and sharing social data
effectively. This aspect of data curation gains further relevance in
the research by \citet{mclure2014data}, which delves into the specific
practices and needs within data curation communities. The practical
implications of data curation are vividly illustrated in the context
of public health and global challenges. \citet{cantor2018facets}
demonstrate the utility of curated open data in evaluating community
health determinants. Furthermore, the COVID-19 pandemic serves as a
real-world example, with \citet{shankar2021data} observing the
critical role of collective data curation efforts in managing and
responding to the crisis. Collectively, these studies not only highlight the 
multifaceted nature of data curation but also emphasize
its indispensable role in enhancing the applicability and value of
open data across various domains.

@@

A note on naming conventions: dataset field names are named 
using ``snake\_case'' where each word in the variable name 
is written in lowercase, and words are separated by 
underscores, e.g. created\_date, complaint\_type, zip\_code, etc..

>>>>

\usepackage[margin = 1in]{geometry}
\usepackage{amsmath}
\allowdisplaybreaks
\usepackage{authblk}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{hyperref}\hypersetup{colorlinks=true, citecolor=blue}
% \usepackage{parskip}
\usepackage{textgreek}
\usepackage{graphicx}
\usepackage{booktabs} % For better looking tables
\usepackage{siunitx} % For alignment of numbers
\usepackage{placeins}
\usepackage{caption}  % For custom captions
\usepackage{float}   % Required for the H specifier
\usepackage{adjustbox}
\usepackage[]{lineno}
\linenumbers*[1]

@@

\usepackage{xcolor}
\newcommand{\dt}[1]{\textcolor{purple}{DT: (#1)}}
\newcommand{\jy}[1]{\textcolor{cyan}{JY: (#1)}}

\let\proglang=\textsf
%% \newcommand{\pkg}[1]{{\fontseries{m}\selectfont #1}}
%% \newcommand\code[2][black]{\textcolor{#1}{\texttt{#2}}}

@@

\begin{abstract}
In the early 21st century, the open data movement began to transform 
societies and governments by promoting transparency,
innovation, and public engagement. New York City (NYC) has been at
the forefront of this movement since the enactment of the Open 
Data Law in 2012, which led to the creation of the NYC Open Data
portal. This portal now hosts 2,700 datasets from 80 city agencies,
serving as a crucial resource for research across various domains, 
including health, urban development, and transportation. The 
success of these initiatives highlights the importance of data 
curation in ensuring the utility and reliability of open datasets.
This paper examines the challenges of open data curation through a
case study of the NYC 311 Service Request (SR), addressing issues 
of data validity, consistency, and curation efficiency. Based on 
insights from this case study, we propose a set of data curation 
principles tailored for government-released open data. These principles 
aim to enhance data management practices and ensure 
the ongoing utility of open data. The paper concludes with 
actionable recommendations for enhancing data curation and outlines
general principles for the effective release of open data.

@@

\jy{cut down to 6-8}
\noindent
\textit{Keywords:}
	311 Service Request Data;
	Data cleansing;
	Data Consistency;
	Data Curation;
	Data Democratization;
	Data Efficiency;
	Data science;
	Data Validity;
	Government Data;
	NYC Open Data;
	Open Data Movement;
	Open data;
	Public Engagement;
	Quality control;
	Research Data Management;
	Smart Cities;
	Transparency;
\end{abstract}

@@

New York City (NYC) has emerged as a leader in the open data movement,
marked by the enactment of the Open Data Law in 2012
\citep{zuiderwijk2014open}. This landmark legislation led to the
creation of the \href{https://opendata.cityofnewyork.us}{NYC Open Data
  portal}, which today hosts an impressive array of 2,700 datasets
from 80 different city agencies. This resource has become invaluable
for researchers across various fields and has significantly enhanced
local government transparency. Popular datasets include information on
restaurant health inspection violations, car crashes, high school and
college enrollment statistics, jail inmate charges, and the location
of city-wide free Internet access points. These datasets have been
applied in civil life in various ways, such as mapping car crashes
involving pedestrians and visualizing high school and college
enrollment trends. Furthermore, they have enabled significant research
across multiple domains, including health \citep{cantor2018facets,
  shankar2021data}, urban development \citep{neves2020impacts}, and
transportation \citep{gerte2019understanding}, aiding in the
understanding and addressing of complex urban challenges.

The civil and research applications of open data critically depend on
its quality, making data curation fundamental in the open data
ecosystem. Among the earliest discussions,
\citet{witt2009constructing} focused on developing data curation
profiles tailored to specific contexts, setting a precedent for
targeted data management strategies. Addressing broader challenges in
data sharing and management, \citet{borgman2012conundrum} highlighted
the complexities of research data distribution, emphasizing the need
for robust strategies. This is complemented by \citet{hart2016ten},
who outlined essential principles for effective data management,
particularly emphasizing meticulous curation practices. In
collaborative data management, \citet{beheshti2019datasynapse}
underscored the significance of cooperative environments for managing
and sharing social data effectively. This aspect gains further
relevance in \citet{mclure2014data}, which delved into the specific
practices and needs within data curation communities. The practical
implications of data curation are vividly illustrated in public health
and global challenges. \citet{cantor2018facets} demonstrated the
utility of curated open data in evaluating community health
determinants. The COVID-19 pandemic served as a real-world example,
with \citet{shankar2021data} observing the critical role of collective
data curation efforts in managing and responding to the
crisis. Collectively, these studies highlight the multifaceted nature
of data curation and emphasize its indispensable role in enhancing the
applicability and value of open data across various domains.

@@

\dt{
A note on naming conventions: dataset field names are named 
using ``snake\_case'' where each word in the variable name 
is written in lowercase, and words are separated by 
underscores, e.g. created\_date, complaint\_type, zip\_code, etc..   
}
